---
title: "Using the ESCSiteData API"
# subtitle: "v0.5"
execute: 
  enabled: true
title-block-banner: "#491656"
css: styles.css
---

## Outline

This notebook details how to call the ESCSiteData API endpoint to retrieve modelled environmental dose values for a given location and climate scenario.

## Import required packages

::: {.panel-tabset}

## Python

```{python import_packages_py}
import requests
import json
import pandas
pandas.options.mode.chained_assignment = None
```

## R

```{r import_packages_r}
library(httr)
library(jsonlite)
library(dplyr)
```

:::

## Compose API request

Establish API POST request URL

::: {.panel-tabset}

## Python

```{python url_py}
base_url = "http://209.97.132.245:8000/ESCSiteData/"
```

## R

```{r url_r}
base_url <- "http://209.97.132.245:8000/ESCSiteData/"
```

:::

Set the longitude (lon) and latitude (lat) in EPSG:3857.

::: {.panel-tabset}

## Python

```{python lonlat_py}
lon = -141102.7542
lat = 6830259.9735
```

## R

```{r lonlat_r}
lon <- -141102.7542
lat <- 6830259.9735
```

:::

Establish the climate scenario/s (climateScenList) and climate years (climateYearList).

* climateScenList:List
    * A list containing any of: ['UKCP09', 'RCP2.6', 'RCP4.5', 'RCP6.0', 'RCP8.5']
* climateYearList:List
    * A list containing any of: ['Baseline', '2010', '2020', '2030', '2040', '2050', '2060', '2070', '2080']

Please note that we are currently integrating UKCP18 RCP scenario data, as such RCP2.6, RCP4.5, RCP6.0, and RCP8.5 are currently placeholders.

::: {.panel-tabset}

## Python

```{python clim_py}
climateScenList = ["RCP2.6"]
climateYearList = ["Baseline", "2010"]
```

## R

```{r clim_r}
climateScenList <- list("RCP2.6")
climateYearList <- list("Baseline", "2010")
```

:::

Compose the request body.

::: {.panel-tabset}

## Python

```{python reqBody_py}
req_body = {"lon": lon, "lat": lat, 
            "climateScenList": climateScenList, "climateYearList": climateYearList}
```

## R

```{r reqBody_r}
req_list <- list("lon" = lon, "lat" = lat, 
                 "climateScenList" = climateScenList, "climateYearList" = climateYearList)

req_body <- jsonlite::toJSON(req_list, auto_unbox = TRUE)
```

:::


## Make the API call

Call the ESCSiteData API.

::: {.panel-tabset}

## Python

```{python apiCall_py}
ESCSiteDataresponse = requests.post(url = base_url, json = req_body)
ESCSiteDataresponse.status_code == 200
```

## R

```{r apiCall_r}
ESCSiteDataresponse <- httr::POST(url = base_url, body = req_body)
ESCSiteDataresponse$status_code == 200
```

:::

Establish a function to deserialize the json response into a nested dictionary or nested list.

::: {.panel-tabset}

## Python

```{python deserializeFunc_py}
def deserialize(responseJson:str) -> dict:

    responseDataFrame = pandas.DataFrame.from_dict(data = json.loads(responseJson))
            
    return responseDataFrame
```

## R

```{r deserializeFunc_r}
deserialize <- function(responseJson){

    responseContent <- httr::content(responseJson)
    responsedf <- jsonlite::fromJSON(responseContent)

    return(responsedf)
}
```

:::

Call the `deserializeResults` function on the ESC API output.

::: {.panel-tabset}

## Python

```{python deserialize_py}
ESCSiteDataresponse_df = deserialize(responseJson = ESCSiteDataresponse.json())
```

## R

```{r deserialize_r}
ESCSiteDataresponse_df <- deserialize(responseJson = ESCSiteDataresponse)
```

:::


## Explore results

The object produced (`ESCSiteDataresponse_df`) contains the environmental dose variables obtained from the modelled data for a given set of climate parameters.
For more information on the underlying method see the [Forest Research Bulletin 124](https://www.forestresearch.gov.uk/publications/archive-an-ecological-site-classification-for-forestry-in-great-britain/) and the [Forest Research Technical Paper 20](https://www.forestresearch.gov.uk/publications/archive-an-ecological-site-classification-for-forestry-in-great-britain-with-special-reference-to-grampian-scotland/).
Column definitions:

* climateScen - climate scenario.
* climateYear - climate year.
* SMR - soil moisture regime.
* SNR - soil nutrient regime.
* AT - accumulated temperature.
* CT - continentality.
* DAMS - detailed aspect method of scoring.
* MD - moisture deficit.

::: {.panel-tabset}

## Python

```{python ESCDoseVals_py}
ESCSiteDataresponse_df
```

## R

```{r ESCDoseVals_r}
ESCSiteDataresponse_df
```

:::